---
title: 'Regrestimate: Predicting Charlotte, NC Home Sale Prices'
author: "Ann (Zi'an) Zhang, Ben Keel"
date: "2022-10-14"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    css:
editor_options: 
  markdown: 
    wrap: 72
---
## Starting Point
<Describing the original data set>

### Acquiring Additional Variables
<Gathering Data Paragraph>


```{r setup, include=FALSE}

# You can set some global options for knitting chunks

knitr::opts_chunk$set(echo = TRUE)

# Load some libraries

library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot) # plot correlation plot
library(corrr)      # another way to plot correlation plot
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance) # to support jtools plots
library(ggpubr)    # plotting R^2 value on ggplot point scatter
library(broom.mixed) # needed for effects plots

# functions and data directory
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#25CB10", "#5AB60C", "#8FA108",   "#C48C04", "#FA7800")
```

### Summary Statistics

<Paragraph>
<Data sorted by internal characteristics, local surroundings, spatial structure>

```{r Summary Statistics, warning = FALSE, message = FALSE}



```


### Correlation Matrix

<Paragraph>

```{r Correlation Matrix, warning = FALSE, message = FALSE}



```

### Home Price Scatterplots

Present 4 home price correlation scatterplots that you think are of interest.

```{r Home Price Scatterplots, warning = FALSE, message = FALSE}



```

### Map of Home Prices and Indicators

<Paragraph>

```{r Map of Home Prices, warning = FALSE, message = FALSE}



```

### Map 1 

<Paragraph>

```{r Indicator Map 1, warning = FALSE, message = FALSE}



```

### Map 2 

<Paragraph>

```{r Indicator Map 2, warning = FALSE, message = FALSE}



```

### Map 3 

<Paragraph>

```{r Indicator Map 3, warning = FALSE, message = FALSE}



```

## Prediction Methods

<Paragraph>

```{r Prediction Methods, warning = FALSE, message = FALSE}



```

### Training Set Results

<Paragraph>

```{r Training Set, warning = FALSE, message = FALSE}



```

### Test Set Results

<Paragraph>

```{r Test Set Results, warning = FALSE, message = FALSE}



```

### CV Results

<Paragraph>

```{r CV Results, warning = FALSE, message = FALSE}



```

### Predicted vs Observation Scatterplot

<Paragraph>

```{r Predicted vs Observation Scatterplot, warning = FALSE, message = FALSE}



```

### Residuals Map

<Paragraph>

```{r Residuals Map, warning = FALSE, message = FALSE}



```

### Predicted Map

<Paragraph>

```{r Predicted Map, warning = FALSE, message = FALSE}



```

## MAPE by Neighborhood

<Paragraph>

```{r MAPE by Neighborhood, warning = FALSE, message = FALSE}



```

### Scatterplot - MAPE by neighborhood mean price

<Paragraph>

```{r MAPE scatterplot, warning = FALSE, message = FALSE}



```

### Split by Census Group

<Paragraph>

```{r Split by Census Group, warning = FALSE, message = FALSE}



```

## Evaluating our Model

### Accuracy

<Paragraph>

### Generalizability

<Paragraph>

## Conclusion