---
title: 'Regrestimate: Predicting Charlotte, NC Home Sale Prices'
author: "Ann (Zi'an) Zhang, Ben Keel"
date: "2022-10-14"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    css:
editor_options: 
  markdown: 
    wrap: 72
---
## Starting Point
<Describing the original data set>

### Acquiring Additional Variables
<Gathering Data Paragraph>


```{r setup, include=FALSE}

# You can set some global options for knitting chunks

knitr::opts_chunk$set(echo = TRUE)

# Load some libraries

library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot) # plot correlation plot
library(corrr)      # another way to plot correlation plot
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance) # to support jtools plots
library(ggpubr)    # plotting R^2 value on ggplot point scatter
library(broom.mixed) # needed for effects plots

# functions and data directory
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#25CB10", "#5AB60C", "#8FA108",   "#C48C04", "#FA7800")
```

### Summary Statistics

<Paragraph>
<Data sorted by internal characteristics, local surroundings, spatial structure>

```{r Summary Statistics, warning = FALSE, message = FALSE}



```


### Correlation Matrix

<Paragraph>

```{r Correlation Matrix, warning = FALSE, message = FALSE}



```

### Home Price Scatterplots

Present 4 home price correlation scatterplots that you think are of interest.

```{r Home Price Scatterplots, warning = FALSE, message = FALSE}



```

### Map of Home Prices and Indicators

<Paragraph>

```{r Map of Home Prices, warning = FALSE, message = FALSE}



```

### Map 1 


```{r Indicator Map 1, warning = FALSE, message = FALSE}



```

### Map 2 

Map 2 of our second most interesting predictor.

```{r Indicator Map 2, warning = FALSE, message = FALSE}



```

### Map 3 

Map 3 of our third most interesting predictor.

```{r Indicator Map 3, warning = FALSE, message = FALSE}



```

## Prediction Methods

<Paragraph>

```{r Prediction Methods, warning = FALSE, message = FALSE}



```

## Results

### Training Set Results

<Paragraph>

```{r Training Set, warning = FALSE, message = FALSE}



```

### Test Set Results

<Paragraph>

```{r Test Set Results, warning = FALSE, message = FALSE}



```

### Cross-Validation Tests & Results

<Paragraph>

```{r CV Results, warning = FALSE, message = FALSE}



```

### Predicted vs Observation Scatterplot

<Paragraph>

```{r Predicted vs Observation Scatterplot, warning = FALSE, message = FALSE}



```

### Residuals Map

<Paragraph>

```{r Residuals Map, warning = FALSE, message = FALSE}



```

### Predicted Map

<Paragraph>

```{r Predicted Map, warning = FALSE, message = FALSE}



```

### MAPE by Neighborhood

<Provide a polished table of mean absolute error and MAPE for a single test set.>

<Paragraph> 

```{r MAPE by Neighborhood, warning = FALSE, message = FALSE}



```

#### Scatterplot - MAPE by neighborhood mean price

<Provide a scatterplot plot of MAPE by neighborhood as a function of mean price by neighborhood.>

<Paragraph>

```{r MAPE scatterplot, warning = FALSE, message = FALSE}



```

### Split by Census Group

<Using tidycensus, split your study area into two groups (perhaps by race or income) and test your model’s generalizability. Is your model generalizable?>

<Paragraph>

```{r Split by Census Group, warning = FALSE, message = FALSE}



```

## Evaluating our Model

Is this an effective model? 
What were some of the more interesting variables?
How much of the variation in prices could you predict?
Describe the more important features? Describe the error in your predictions? 
According to your maps, could you account the spatial variation in prices?  
Where did the model predict particularly well? Poorly? Why do you think this might be?

### Accuracy


### Generalizability

<Paragraph>

## Conclusion