---
title: 'Regrestimate: Predicting Charlotte, NC Home Sale Prices'
author: "Ann (Zi'an) Zhang, Ben Keel"
date: "2022-10-14"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    css:
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

This project uses the Hedonic Model and OLS regression for predicting
home price in Mecklenberg County, NC. We use existing 46,183 cases to
build the model -- 60% for training the model and 40% for testing the
efficacy of the model. Once we produce a model with satisfactory
efficacy, we then use this model to make prediction for 100 home units.

To generate a model with stronger predictive power, we incorporated more
valid internal (structural) and external (neighborhoods, amenities,
etc.) factors that may affect home price. The selection and engineering
of those factors were executed with careful consideration of Mecklenberg
County (including the City of Charlotte)'s local circumstances. After
consulting local residents of Charlotte, we included factors like
schools, crime, and area of parcel since schooling, safety, and having a
yard are among their primary concerns when choosing houses.

(Brief Results)

# Data Source

The original data set contains basic information about the homes,
including price, structural factors (e.g. number of stories), and
ownership information. It consists of a modelling set of 46081
observations with existing price data and a challenge set of 100
observations, of which price is to be predicted using the model we
generate.

In addition to the original data set provided to us, we selected data
sets that are deemed valid factors for predicting housing price from the
City of Charlotte Open Data Portal (<https://data.charlottenc.gov/>) and
Charlotte/Mecklenburg Quality of Life Explorer, an online portal created
by Mecklenburg County, the City of Charlotte, and UNC Charlotte
(<https://mcmap.org/qol/#43/>).

**Note:** information about neighborhood in the county are taken from
the Quality of Life Explorer. Mecklenburg County currently utilizes the
system of "neighborhood profile area (NPA)," which is generated from the
old "neighborhood statistic area (NSA)." Instead of neighborhood names,
NPA is encoded in ID-numbers. There are 458 NPAs in the Mecklenburg
County.

# Data Analysis

This project requires several R packages to be loaded, including but not
limited to 'FNN' package for calculating the distance to the nearest
shop/hospital/pharmacy ('K-nearest-neighbor') and 'stargazer' and
'kableExtra' for constructing tables.

```{r setup, include=FALSE}

# You can set some global options for knitting chunks

knitr::opts_chunk$set(echo = TRUE)

# Load some libraries

library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot) # plot correlation plot
library(corrr)      # another way to plot correlation plot
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance) # to support jtools plots
library(ggpubr)    # plotting R^2 value on ggplot point scatter
library(broom.mixed) # needed for effects plots
library(stargazer)

g <-glimpse

options(scipen=999)

# functions and data directory
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

library(tidycensus)
census_api_key("3c9540be1434ac4b38e6e55d60e8ee95909f2254", overwrite = TRUE)

palette5 <- c("#25CB10", "#5AB60C", "#8FA108",   "#C48C04", "#FA7800")
Pink <- c("#ffffff","#f9f4f4","#f0e4e4", "#e7d3d3", "#dec3c3")
Blue <- c("#f8fbff", "#eaf4ff", "#d6eaff", "#add6ff", "#84c1ff")
Violet <- c("#ffffff", "#f7f7f7", "#dfe3ee", "#8b9dc3",  "#58668b")
Green <- c("#e8f4ea", "#e0f0e3", "#d2e7d6","#c8e1cc", "#b8d8be")
Purple <- c("#f3e0f7", "#e4c7f1", "#d1afe8", "#b998dd", "#9f82ce")
Teal <- c("#d1eeea", "#a8dbd9", "#85c4c9", "#68abb8", "#4f90a6")

```

```{r}

#original dataset
Charlotte.nhoods <- st_read(file.path("https://bennkeel.github.io/Zestimate/Area(1).geojson"))  %>%
  st_transform('ESRI:103501')%>%
  dplyr::select(-X2020)%>%
  rename(npa = id)

g(Charlotte.nhoods)

Charlotte <-
  st_read("https://raw.githubusercontent.com/mafichman/MUSA_508_Lab/main/Midterm/data/2022/studentData.geojson") %>%
  st_transform('ESRI:103501')

```

```{r Variables We Liked}

Charlotte.Clean <-
  Charlotte %>%
  dplyr::select("pid", "nc_pin", "municipali", "yearbuilt", "heatedarea", "price", "storyheigh", "heatedfuel", "actype", "extwall", "foundation", "numfirepla", "fireplaces", "bldggrade", "fullbaths", "halfbaths", "bedrooms", "toPredict", "landusecod", "shape_Area", "musaID") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501') %>%
  filter(price <= 10000000)

Charlotte.Clean$fireplaces[is.na(Charlotte.Clean$fireplaces)] <- "FP0"

Charlotte.Clean <-
  st_join(Charlotte.Clean, Charlotte.nhoods, join=st_intersects)


```

## Additional Variables & Factor Engineering

After careful consideration of potential factors that impact the housing
price, we extracted data of parks, schools, medical facilities, shops
(including grocery stores and shopping malls), pharmacies, homicide
(representing high-profile crime), and incidents (low-profile crime)
from the open data portal. They become the independent variables in our
model.

We then conducted factor engineering, namely, transforming existing data
into a factor by establishing relationship between the independent
variables with the dependent variable (home price) in two ways. Here, we
use "k-nearest-neighbor" to calculated the distance from each home to
the closest, 2nd closest, and 3rd closet shop/pharmacy, etc. and
"aggregation within buffer" to calculated the number of shop/pharmacy,
etc. within 1 mile range from each home.

```{r}

#Park location information
park <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/735a6bce6306442face38657b50fc7b7_10/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_transform('ESRI:103501') 
  
park.buffer <-
  park %>%
  dplyr::select(geometry) %>%
  st_transform('ESRI:103501') %>%
    na.omit() 

Charlotte.Clean$park.Buffer <- 
    Charlotte.Clean$geometry %>% 
    st_buffer(5280) %>% 
    aggregate(mutate(park.buffer, counter = 1),., sum) %>%
    pull(counter)

Charlotte.Clean$park.Buffer[is.na(Charlotte.Clean$park.Buffer)] <- 0


#School location information
school <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/04f2ea0b58774ee7b2e525816cbbc0bb_1/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501') %>%
  filter(!st_is_empty(.))

school.buffer <-
  school %>%
  dplyr::select(geometry) %>%
  st_transform('ESRI:103501') %>%
    na.omit()

Charlotte.Clean$school.Buffer <- 
    Charlotte.Clean$geometry %>% 
    st_buffer(5280) %>% 
    aggregate(mutate(school.buffer, counter = 1),., sum) %>%
    pull(counter)

Charlotte.Clean$school.Buffer[is.na(Charlotte.Clean$school.Buffer)] <- 0


#Hospital Proximity
medical <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/320dbc7d1ef944f5bf7c5e21b018b678_4/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501')

med.buffer <-
  medical %>%
  dplyr::select(geometry) %>%
  st_transform('ESRI:103501') %>%
    na.omit() 

Charlotte.Clean$med.Buffer <- 
    Charlotte.Clean$geometry %>% 
    st_buffer(5280) %>% 
    aggregate(mutate(med.buffer, counter = 1),., sum) %>%
    pull(counter)

Charlotte.Clean$med.Buffer[is.na(Charlotte.Clean$med.Buffer)] <- 0
  

#Food Access
grocery <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/93e54082dde0418d836a57f2fc12879f_7/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_transform('ESRI:103501') %>%
  dplyr::select("OBJECTID", "geometry")

shoppingmall <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/58487298f4ee455e84e236b5db43195d_11/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501') %>%
  dplyr::select("OBJECTID", "geometry")

shopping <-
  rbind(grocery, shoppingmall)

Charlotte.Clean <-
  Charlotte.Clean %>% 
    mutate(
      shop_nn1 = nn_function(st_coordinates(Charlotte.Clean), 
                              st_coordinates(shopping), k = 1))

#Health Access
pharmacies <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/dcfdb72bc9c045a0b0945da79a966841_3/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501') %>%
  filter(!st_is_empty(.))

Charlotte.Clean <-
  Charlotte.Clean %>% 
    mutate(pharm_nn1 = nn_function(st_coordinates(Charlotte.Clean), 
                              st_coordinates(pharmacies), k = 1))

#High-profile crime
homicide <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/b5b21fcd2ad24de9ba7a13093648f5e9_0/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_transform('ESRI:103501')

homicide.buffer <-
  homicide %>%
  dplyr::select(LATITUDE_PUBLIC, LONGITUDE_PUBLIC) %>%
  st_as_sf(coords = c("LONGITUDE_PUBLIC", "LATITUDE_PUBLIC"), crs = 4326) %>%
  st_transform('ESRI:103501') %>%
    na.omit() 

Charlotte.Clean$homi.Buffer <- 
    Charlotte.Clean$geometry %>% 
    st_buffer(5280) %>% 
    aggregate(mutate(homicide.buffer, counter = 1),., sum) %>%
    pull(counter)

Charlotte.Clean$homi.Buffer[is.na(Charlotte.Clean$homi.Buffer)] <- 0


#Low-profile Crime
incidents <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/d22200cd879248fcb2258e6840bd6726_0/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501') %>%
  filter(!st_is_empty(.))

Charlotte.Clean <-
  Charlotte.Clean %>% 
    mutate(inci_nn1 = nn_function(st_coordinates(Charlotte.Clean), 
                              st_coordinates(incidents), k = 1))


```

### Separating Modeling vs. Predicting Sets

Since the original data set contains both the data used for making the
prediction model and the data to be predicted by the model, we separate
the original data set into "modelling" (for training model) and
"challenge" (to be predicted) sets.

```{r warning = FALSE, message = FALSE}

Charlotte.Modelling <-
  Charlotte.Clean %>%
  filter(toPredict=="MODELLING") 

Charlotte.Challenge <-
  Charlotte.Clean %>%
  filter(toPredict=="CHALLENGE")


```

## Summary Statistics

To better understand the variables that we considered as potentially
valid factors for predicting home prices, the following summary
statistic tables are created. The variables are split into the internal
characteristics, which are the inherent structural characteristics that
come with the house itself, and external factors, which includes local
surroundings, distance to amenities and civic services, and spatial
structures. Note that the tables only support summarizing numeric data,
so categorical variables such as 'fireplaces' (type of fireplaces) are
not included in the following tables.

```{r Summary Statistics, warning = FALSE, message = FALSE}

Charlotte.Summary <-
  Charlotte.Modelling %>%
  dplyr::select(-shape_Leng, -musaID, -id, -shape_Area.y) %>%
  st_drop_geometry %>%
  na.omit()

stargazer(Charlotte.Summary, type="text", keep = c("price", "yearbuilt", "heatedarea", "storyheigh", "heatedfuel", "actype", "extwall", "foundation", "numfirepla", "fireplaces", "bldggrade", "fullbaths", "halfbaths", "bedrooms"), title = "Table 1.1. Internal Characteristics Summary Statistics" )

```

```{r}
stargazer(Charlotte.Summary, type="text", omit = c("price", "yearbuilt", "heatedarea", "storyheigh", "heatedfuel", "actype", "extwall", "foundation", "numfirepla", "fireplaces", "bldggrade", "fullbaths", "halfbaths", "bedrooms"), title = "Table 1.2. External Characteristics Summary Statistics")
```

## Correlation Matrix

```{r Correlation Matrix, warning = FALSE, message = FALSE, fig.width=6}

numericVars <- 
  select_if(st_drop_geometry(Charlotte.Modelling), is.numeric) %>% na.omit()

ggcorrplot(
  round(cor(numericVars), 1), 
  p.mat = cor_pmat(numericVars),
  colors = c("#B0E0E6", "white", "#B272A6"),
  type="lower",
  insig = "blank") +  
    labs(title = "Correlation Across Numeric Variables", caption = "Fig. 2") 

```

This correlation matrix (Fig. 2) shows correlation across numeric
variables, including both internal structural and external neighborhood
characteristics. Higher correlation between any of the two variables
(other than price) suggest the possibility of col-linearity. As a
result, col-linear variables should be reduced to one that encapsulate
the effect of this factor. For example, since "shop_nn1," "shop_nn2,"
and "shop_nn3" (distance to the 1st/2nd/3rd closest shop) are all
col-linear with each other, we chose to only incorporate "shop_nn1"
(distance to the closest shop) into our regression model for home price
prediction and discard the other two.

## Home Price Scatter Plots

We selected four variables -- distances to the nearby homicide (Fig.
3.1), incident (Fig. 3.2), shops (Fig. 3.3), and schools (Fig. 3.4) to
look at their relationship with home prices via scatter plots. We chose
to look at the nearest 5 homicide and incidents and 3 shops/schools
because we assume residents wouldn't care too much about the 4th or 5th
closest shops/schools -- they would just go to closer ones.

```{r Home Price Scatterplots 1, warning = FALSE, message = FALSE, fig.width=6}

#homicide

Charlotte.Modelling %>%  
  st_drop_geometry() %>%
  filter(price <= 1000000) %>%
  dplyr::select(price, starts_with("homo_")) %>%
  gather(Variable, Value, -price) %>%
  ggplot(aes(Value, price)) +
      geom_point(size = .1, color = "light grey") +
      geom_smooth(method = "lm", se=F, color = "black") +
      facet_wrap(~Variable, nrow = 1, scales = "free") +
      labs(title = "Home Price & Distance to the (1st/2nd/3rd) Nearest Homicide", subtitle = "Mecklenburg County, NC", caption = "Fig. 3.1") +
      xlab("Distance (in ft.)") +
      ylab("Home Price") +
  theme(strip.text.x = element_text(size = 3)) +
  plotTheme()



```

```{r warning = FALSE, message = FALSE, fig.width=6}
#inci_nn1


Charlotte.Modelling %>%  
  st_drop_geometry() %>%
  filter(price <= 1000000) %>%
  dplyr::select(price, starts_with("inci_")) %>%
  gather(Variable, Value, -price) %>%
  ggplot(aes(Value, price)) +
      geom_point(size = .1, color = "#B0C4DE") +
      geom_smooth(method = "lm", se=F, color = "black") +
      facet_wrap(~Variable, nrow = 1, scales = "free") +
      labs(title = "Home Price & Distance to the (1st/2nd/3rd) Nearest Incident", subtitle = "Mecklenburg County, NC", caption = "Fig. 3.2") +
      xlab("Distance (in ft.)") +
      ylab("Home Price") +
  theme(strip.text.x = element_text(size = 3)) +
  plotTheme()

```

From the two scatter plots above, we can see a clear relationship that
the further a home is from a homicide and/or incident, the higher its
price may get. This aligns with the common understanding of people
paying more for homes in 'safer areas' (with less incidents and
homicide).

```{r warning = FALSE, message = FALSE, fig.width=6}

#shop_nn

Charlotte.Modelling %>%  
  st_drop_geometry() %>%
  filter(price <= 1000000) %>%
  dplyr::select(price, starts_with("shop_")) %>%
  gather(Variable, Value, -price) %>%
  ggplot(aes(Value, price)) +
      geom_point(size = .1, color = "#D8BFD8") +
      geom_smooth(method = "lm", se=F, color = "dark grey") +
      facet_wrap(~Variable, nrow = 1, scales = "free") +
      labs(title = "Home Price & Distance to the (1st/2nd/3rd) Nearest Shop", subtitle = "Mecklenburg County, NC", caption = "Fig. 3.3") +
      xlab("Distance (in ft.)") +
      ylab("Home Price") +
  theme(strip.text.x = element_text(size = 3)) +
  plotTheme()
 
```

```{r  warning = FALSE, message = FALSE, fig.width=6}

#school_nn
Charlotte.Modelling %>%  
  st_drop_geometry() %>%
  filter(price <= 1000000) %>%
  dplyr::select(price, starts_with("school_")) %>%
  gather(Variable, Value, -price) %>%
  ggplot(aes(Value, price)) +
      geom_point(size = .1, color = "#B4D7BF") +
      geom_smooth(method = "lm", se=F, color = "dark grey") +
      facet_wrap(~Variable, nrow = 1, scales = "free") +
      labs(title = "Home Price & Distance to the (1st/2nd/3rd) School", subtitle = "Mecklenburg County, NC", caption = "Fig. 3.4") +
      xlab("Distance (in ft.)") +
      ylab("Home Price") +
  theme(strip.text.x = element_text(size = 3)) +
  plotTheme()


```

Fig. 3.3 and Fig. 3.4, on the other hand, provides interesting insights
-- the further homes are from closest shops or schools, the higher their
price may get. This contradicts the expectations that convenience (to
shops and schools) comes at a price. This may be cause by a possible
confounding factor, such as closer to shops and schools might mean it's
a more densely populated area, although it remains uncertain.

### Park Buffer - Interesting Observation

```{r}

Charlotte.Modelling %>%  
  st_drop_geometry() %>%
  filter(price <= 1000000) %>%
  dplyr::select(price, starts_with("park_")) %>%
  gather(Variable, Value, -price) %>%
  ggplot(aes(Value, price)) +
      geom_point(size = .1, color = "light grey") +
      geom_smooth(method = "lm", se=F, color = "black") +
      facet_wrap(~Variable, nrow = 1, scales = "free") +
      labs(title = "Home Price & Distance to Closest Parks", subtitle = "Mecklenburg County, NC", caption = "Fig. 5.4") +
      xlab("Distance (in ft.)") +
      ylab("Home Price") +
  theme(strip.text.x = element_text(size = 3)) +
  plotTheme()

```

Another interesting observation is plotted in Fig. 5.4. If we look at
the relationship between home prices and their distance to the closest
parks, we see variation across the closest, 2nd closest, and 3rd closest
parks. The scatter plot suggest that the further the 'closest' park
lays, the home price goes down. However, the 2nd closest park doesn't
have a noticeable effect on home price, and the 3rd closest have the
opposite effect --- the further the 3rd closest park is, the home price
goes up.

This phenomenon may be caused the fact that if a home is located closely
to multiple parks, it may lay in a rather rural / remote area that are
too far from civic services, making it inconvenient and unappealing.

## Map of Home Prices and 3 Factors

We first mapped the home prices (Fig 5.1) to see the distribution of
home prices in existing 'Modelling' data set. Then, we mapped three
factors of interest - school buffers (which indicate the number of
schools within 1 mile buffer from each home) (Fig. 5.2), distance to
nearest shop (Fig. 5.3), and land use codes (Fig 5.4).

By comparing the latter 3 maps with the home price map, we can see that
homes are more expensive if they have more schools within 1 mile range,
especially concentrated in the southern area; home price is less
associated with distance to the nearest shops, since they are more
spread across the county. The land use code map brings us insight of the
distribution of different types of homes. From the map we can see that
the majority are single-family residents, with some categorized as near
"waterfront" or "rural acreage."

```{r Map of Home Prices, warning = FALSE, message = FALSE, fig.width=6}

ggplot() + 
  geom_sf(data = st_union(Charlotte.nhoods), fill = "#000000") +
  geom_sf(data = Charlotte.Modelling, aes(color = q5(price)), show.legend = "point", size = .1) +
    scale_colour_manual(values = Blue, labels=qBr(Charlotte.Modelling,"price"), name="Home Price") +
    labs(title = "Home Price (Dependent Variable)", subtitle = "Mecklenburg County, NC", caption = "Fig. 6.1") +
                          mapTheme()

```

### Map 1 - School Buffer

```{r Indicator Map 1, warning = FALSE, message = FALSE}

ggplot() + 
  geom_sf(data = st_union(Charlotte.nhoods), fill = "#000000") +
  geom_sf(data = Charlotte.Modelling, aes(color = q5(school.Buffer)), show.legend = "point", size = .1) +
    scale_colour_manual(values = Pink, labels=qBr(Charlotte.Modelling,"school.Buffer"), name="School Buffer") +
    labs(title = "Number of Schools Within 1 Mile Buffer", subtitle = "Mecklenburg County, NC", aes(color= "pink"), caption = "Fig. 6.2") +
                          mapTheme()

```

### Map 2 - Shop nn1

```{r Indicator Map 2, warning = FALSE, message = FALSE}

ggplot() + 
  geom_sf(data = st_union(Charlotte.nhoods), fill = "#000000") +
  geom_sf(data = Charlotte.Modelling, aes(color = q5(shop_nn1)), show.legend = "point", size = .1) +
    scale_colour_manual(values = Green, labels=qBr(Charlotte.Modelling,"shop_nn1"), name="Home Price") +
    labs(title = "Distance to the Nearest Shop (Independent Variable 2)", subtitle = "Mecklenburg County, NC", caption = "Fig. 6.3") +
                          mapTheme()



```

### Map 3 land use code

```{r Indicator Map 3, warning = FALSE, message = FALSE}

ggplot() + 
  geom_sf(data = st_union(Charlotte.nhoods), fill = "#203354") +
  geom_sf(data = Charlotte.Modelling, aes(color=landusecod), show.legend = "point", size = .1) +
    labs(title = "Land Use Code (Inependent Variable)", subtitle = "Mecklenburg County, NC", caption = "Fig. 6.4") +
                          mapTheme()

```

# Prediction Methods

<Paragraph> OLS Regression, cross-validation (k-fold) Splitting the
modelling set into training (60%) and testing (40%)

```{r Prediction Methods, warning = FALSE, message = FALSE}

inTrain <- createDataPartition(
              y = paste(Charlotte.Modelling$storyheigh, Charlotte.Modelling$actype, Charlotte.Modelling$aheatingty, Charlotte.Modelling$heatedfuel, Charlotte.Modelling$extwall, Charlotte.Modelling$foundation, Charlotte.Modelling$numfirepla, Charlotte.Modelling$fireplaces, Charlotte.Modelling$bldggrade, Charlotte.Modelling$fullbaths, Charlotte.Modelling$halfbaths, Charlotte.Modelling$bedrooms, Charlotte.Modelling$municipali, Charlotte.Modelling$landusecod), 
              p = .60, list = FALSE)
Charlotte.training <- Charlotte.Modelling[inTrain,] 
Charlotte.test <- Charlotte.Modelling[-inTrain,]   

```

## Results

### Training Set Results

<Paragraph>

```{r Training Set, warning = FALSE, message = FALSE, results=hide}


reg.training <- 
  lm(price ~ ., data = as.data.frame(Charlotte.training) %>% 
  dplyr::select(price, yearbuilt, heatedarea, storyheigh, heatedfuel, actype, extwall, foundation, numfirepla, fireplaces, bldggrade, fullbaths, halfbaths, bedrooms, pharm_nn1, inci_nn1, homi.Buffer, school.Buffer, med.Buffer, municipali, landusecod, park.Buffer, shop_nn1, shape_Area))

summary(reg.training)

#Quick Summary (raw info)
reg.training%>%
  tidy()%>%
  kable(
    caption = "<center><strong>Regression Variables</strong></center>",
    escape= FALSE,
    format="html")%>%
  kable_styling(bootstrap_options = "striped")

###Does not have r-squared, etc at bottom.

```

### Test Set Measurements

<Paragraph>

```{r Test Set Results, warning = FALSE, message = FALSE}

Charlotte.test <-
  Charlotte.test %>%
  mutate(Regression = "Baseline Regression",
         Price.Predict = predict(reg.training, Charlotte.test),
         Price.Error = Price.Predict - price,
         Price.AbsError = abs(Price.Predict - price),
         Price.APE = (abs(Price.Predict - price)) / Price.Predict) %>%
  filter(price < 10000000) %>%
  na.omit(Charlotte.test)

#Kable table for summary of Mean Absolute Error and MAPE(%)

st_drop_geometry(Charlotte.test)%>%
  dplyr::select(Price.AbsError, Price.APE)%>%
  summarize_at(vars(Price.AbsError, Price.APE), list(mean))%>%
  rename(MeanAbsoluteError = Price.AbsError,
         MAPE = Price.APE)%>%
  kable(
    caption = "<center><strong>Average Regression Errors by Value and Percent</strong></center>",
    escape= FALSE,
    format="html",
    row.names = FALSE,
    align="l")%>%
  kable_styling()

```

### Cross-Validation Tests & Results

<Paragraph>

```{r CV}

fitControl <- trainControl(method = "cv", number = 100)
set.seed(201)

reg.cv <- 
  train(price ~ ., data = st_drop_geometry(Charlotte.Clean) %>% 
                              dplyr::select(price, yearbuilt, heatedarea, storyheigh, heatedfuel,
                                            actype, extwall, foundation, numfirepla, fireplaces,
                                            bldggrade, fullbaths, halfbaths, bedrooms, municipali,
                                            landusecod),
        method = "lm", trControl = fitControl, na.action = na.pass)

summary.cv <- reg.cv$resample[1:100,]

data.frame(Standard_Deviation = c(sd(summary.cv$MAE)), 
          Mean = c(mean(summary.cv$MAE)))%>%
    kable(
    caption = "<center><strong>Cross Validation: Mean Absolute Error of 100 Samples</strong></center>",
    escape= FALSE,
    format="html",
    row.names = FALSE, 
    align="l")%>%
  kable_styling()


```

#### K-fold Cross Validation

```{r CV Results, warning = FALSE, message = FALSE}

ggplot(summary.cv, aes(x=MAE)) +
  geom_histogram(color="blue", fill = "blue")+
  labs(title = "Distribution of MAE")+
  plotTheme()

```

#### Is the model generalizable to new data?

<Talk about it here>

### Predicted vs Observation Scatter plot

<Paragraph>

```{r Predicted vs Observation Scatterplot, warning = FALSE, message = FALSE}

g(Charlotte.test)

summary.fitTest <- Charlotte.test%>%
  dplyr::select(price, Price.Predict)%>%
  rename(Observed_Price = price,
         Predicted_Price = Price.Predict)

ggscatter(summary.fitTest,
          x = "Observed_Price",
          y = "Predicted_Price") +
  geom_smooth(color = "green", size = 2, method = "lm")+
  geom_abline(color = "orange", size = 2) +
  labs(title = "Predicted Prices as a Function of Observed Prices", 
       subtitle = "The orange line is a perfect prediction, \nthe green line is our prediction")+
  stat_cor(label.y = 4000000)


```

### Residuals Map

<Paragraph>

```{r Residuals Map, warning = FALSE, message = FALSE}

ggplot()+
  geom_sf(data=Charlotte.nhoods, fill="gray", color="black")+
  geom_sf(data=Charlotte.test, aes(color=q5(Price.AbsError)), size = 0.05)+
  scale_color_manual(values = palette5,
                     labels = qBr(Charlotte.test, "Price.AbsError"),
                     name= "Price Error\n(Quintile Breaks)")+
  labs(title="Residual Error of Price Estimation", subtitle = "Mecklenburg County, NC")+
  mapTheme()

```

#### Spatial Lag

```{r}

#Spatial lag of Errors
coords <- st_coordinates(Charlotte.Clean) 

neighborList <- knn2nb(knearneigh(coords, 5))

spatialWeights <- nb2listw(neighborList, style="W")

Charlotte.Clean$lagPrice <- lag.listw(spatialWeights, Charlotte.Clean$price)

#Spatial lag of Errors
coords.test <- st_coordinates(Charlotte.test) 

neighborList.test <- knn2nb(knearneigh(coords, 5))

spatialWeights.test <- nb2listw(neighborList, style="W")

Charlotte.test$lagPrice.test <- lag.listw(spatialWeights, Charlotte.test$Price.Error)

#Sales Price as Function of Spatial Lag of Price

lagClean <- ggscatter(Charlotte.Clean,
          x = "lagPrice",
          y = "price",
          xlab ="Spatial Lag of Price (Mean price 5 nearest neighbors)",
          ylab = "Observed Sale Price") +
  geom_smooth(color = "green", size = 2, method = "lm")+
  labs(title = "Sales Price as Function of Spatial Lag of Price")

#Sales Price as Function of Spatial Lag of Error
lagError <- 
  ggscatter(Charlotte.test,
          x = "lagPrice",
          y = "price",
          xlab ="Spatial Lag of Errors (Mean error 5 nearest neighbors)",
          ylab = "Observed Sale Price") +
  geom_smooth(color = "green", size = 2, method = "lm")+
  labs(title = "Sales Price as Function of Spatial Lag of Price") 

grid.arrange(lagClean, lagError)


```

#### Moran's I test

```{r message=FALSE, warning=FALSE}

moranTest <- moran.mc(Charlotte.test$Price.Error, 
                      spatialWeights.test, nsim = 999)

ggplot(as.data.frame(moranTest$res[c(1:999)]), aes(moranTest$res[c(1:999)])) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(aes(xintercept = moranTest$statistic), colour = "#FA7800",size=1) +
  scale_x_continuous(limits = c(-1, 1)) +
  labs(title="Observed and permuted Moran's I",
       subtitle= "Observed Moran's I in orange",
       x="Moran's I",
       y="Count") +
  plotTheme()

```

### Predicted Map

<Paragraph>

```{r Predicted Map, warning = FALSE, message = FALSE}

predictAll <- lm(price ~ ., data = as.data.frame(Charlotte.Clean) %>% 
                   dplyr::select(price, yearbuilt, heatedarea, storyheigh, heatedfuel, 
                                 actype, extwall, foundation, numfirepla, fireplaces, 
                                 bldggrade, fullbaths, halfbaths, bedrooms, municipali, 
                                 landusecod))

Charlotte.PredictAll <- Charlotte.Clean %>%
  mutate(Price.Predict = predict(predictAll, Charlotte.Clean))

ggplot()+
  geom_sf(data=st_union(Charlotte.nhoods), fill="gray", color="black")+
  geom_sf(data=Charlotte.PredictAll, aes(color=q5(Price.Predict)), size = 0.05)+
  scale_color_manual(values = palette5,
                     labels = qBr(Charlotte.test, "Price.Predict"),
                     name= "Prices ($)\n(Quintile Breaks)")+
  labs(title="Home Price Prediction", subtitle = "Mecklenburg County, NC")+
  mapTheme()


stargazer(reg.training, type="text", keep = )

```

### MAPE by Neighborhood

\<Provide a polished table of mean absolute error and MAPE for a single
test set.\>

<Paragraph>

```{r MAPE by Neighborhood, warning = FALSE, message = FALSE}

#Prediction based on neighborhood (From Text)

left_join(
  st_drop_geometry(Charlotte.test) %>%
    group_by(npa) %>%
    summarize(meanPrice = mean(price, na.rm = T)),
  mutate(Charlotte.test, predict.fe = 
                        predict(lm(price ~ npa, data = Charlotte.test), 
                        Charlotte.test)) %>%
    st_drop_geometry %>%
    group_by(npa) %>%
      summarize(meanPrediction = mean(predict.fe))) %>%
      kable() %>% kable_styling()

#MAPE based on neighborhood

Charlotte.npaMAPE <- st_drop_geometry(Charlotte.test) %>%
    group_by(npa) %>%
    summarize(MAPE = mean(Price.APE, na.rm = T), 
              MeanPrice = mean(price, na.rm=T))

Charlotte.npaMAPE%>%
    kable(caption = "<strong>Mean Absolute Percent Error (MAPE) by Neighborhood Profile Areas     (NPA's)</strong>",
    escape= FALSE,
    format="html",
    align = "l") %>% 
    kable_styling(bootstrap_options = "striped")

```

#### Map - MAPE by neighborhood mean price

\<Provide a scatter plot of MAPE by neighborhood as a function of mean
price by neighborhood.\>

<Paragraph>

```{r MAPE map by neighborhood, warning = FALSE, message = FALSE}

g(Charlotte.test)

Charlotte.npaMAPEmap <- left_join(Charlotte.nhoods, Charlotte.npaMAPE, by="npa")

 ggplot()+
  geom_sf(data=Charlotte.npaMAPEmap, aes(fill=q5(MAPE)))+
  scale_fill_manual(values = Blue,
                     labels = qBr(Charlotte.npaMAPEmap, "MAPE"),
                     name= "MAPE\n(Quintile Breaks)")+
  labs(title="MAPE by Neighborhood", subtitle = "Mecklenburg County, NC")+
  mapTheme()

```

### Accounting for Neighborhood in the Model

```{r Neighborhood Model, warning = FALSE, message = FALSE, results = "hide"}

inTrain.nhood <- createDataPartition(
              y = paste(Charlotte.Modelling$storyheigh, Charlotte.Modelling$actype, Charlotte.Modelling$aheatingty, Charlotte.Modelling$heatedfuel, Charlotte.Modelling$extwall, Charlotte.Modelling$foundation, Charlotte.Modelling$numfirepla, Charlotte.Modelling$fireplaces, Charlotte.Modelling$bldggrade, Charlotte.Modelling$fullbaths, Charlotte.Modelling$halfbaths, Charlotte.Modelling$bedrooms, Charlotte.Modelling$municipali, Charlotte.Modelling$landusecod, Charlotte.Modelling$npa), 
              p = .60, list = FALSE)
Charlotte.training.nhood <- Charlotte.Modelling[inTrain.nhood,] 
Charlotte.test.nhood <- Charlotte.Modelling[-inTrain.nhood,]  


reg.nhood <- lm(price ~ ., data = as.data.frame(Charlotte.training.nhood) %>% 
                                 dplyr::select(price, yearbuilt, heatedarea, 
                                               storyheigh, heatedfuel, actype, extwall, 
                                               foundation, numfirepla, fireplaces, 
                                               bldggrade, fullbaths, halfbaths, bedrooms, 
                                               pharm_nn1, inci_nn1, 
                                               homi.Buffer, school.Buffer, med.Buffer, municipali, landusecod, shape_Area, npa))

summary(reg.nhood)

Charlotte.test.nhoods <-
  Charlotte.test.nhood %>%
  mutate(Regression = "Neighborhood Effects",
         Price.Predict = predict(reg.nhood, Charlotte.test.nhood),
         Price.Error = Price.Predict - price,
         Price.AbsError = abs(Price.Predict - price),
         Price.APE = (abs(Price.Predict - price)) / Price.Predict) %>%
  filter(price < 10000000) 

```

#### Scatter Plot - MAPE by neighborhood mean price

\<Provide a scatter plot plot of MAPE by neighborhood as a function of
mean price by neighborhood.\>

<Paragraph>

```{r MAPE scatterplot, warning = FALSE, message = FALSE}

ggscatter(Charlotte.npaMAPE,
          x = "MeanPrice",
          y = "MAPE",
          xlab ="Average Price of Homes in the Neighborhood)",
          ylab = "Mean Absolute Percent Error (MAPE)") +
  labs(title = "Neighborhoods: MAPE as a function of mean price") 

```

### Split by Census Group

\<Using tidycensus, split your study area into two groups (perhaps by
race or income) and test your model's generalizability. Is your model
generalizable?\>

<Paragraph>

```{r Split by Census Group Map, warning = FALSE, message = FALSE}

#Majority Above 100% AMI for household of 2-3

tracts20 <- 
  get_acs(geography = "tract", 
          variables = c("B25010_001", "B19001_001",
                        "B19001_013", "B19001_014", 
                        "B19001_015", "B19001_016", 
                        "B19001_017"), 
          year=2020, state=37,
          county=119, geometry=TRUE, output = 'wide') %>% 
  st_transform('ESRI:103501') %>%
  rename(MeanHHSize = B25010_001E,
         totalHH = B19001_001E,
         HHIncome75to99 = B19001_013E,
         HHIncome100to124 = B19001_014E,
         HHIncome125to149 = B19001_015E,
         HHIncome150to199 = B19001_016E,
         HHIncome200up = B19001_017E
         ) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(HHAbove75 = HHIncome75to99 + HHIncome100to124 + HHIncome125to149 +
         HHIncome150to199 + HHIncome200up, 
         pctAbove75 = HHAbove75/totalHH,
         above75 = ifelse(pctAbove75 > .5, "High Income", "Low Income")) %>%
  dplyr::select(-HHIncome75to99, -HHIncome100to124, -HHIncome125to149,
                -HHIncome150to199, -HHIncome200up)

g(tracts20)

median(tracts20$MeanHHSize, na.rm=TRUE)

ggplot() + geom_sf(data = na.omit(tracts20), aes(fill = above75)) +
  scale_fill_manual(values = c("#25CB10", "#FA7800"), name="Income Level of tract") +
  labs(title = "Annual Income above $75k") +
  mapTheme() + theme(legend.position="bottom")

```

```{r Census Group Analysis Chart}
#neighborhood regression model (we'll substitute with npa)
reg.nhood <- lm(SalePrice ~ ., data = as.data.frame(boston.training) %>% 
                                 dplyr::select(Name, SalePrice, LivingArea, 
                                               Style, GROSS_AREA, NUM_FLOORS.cat,
                                               R_BDRMS, R_FULL_BTH, R_HALF_BTH, 
                                               R_KITCH, R_AC, R_FPLACE,crimes.Buffer))

#creating df with neighborhood adjustments
boston.test.nhood <-
  boston.test %>%
  mutate(Regression = "Neighborhood Effects",
         SalePrice.Predict = predict(reg.nhood, boston.test),
         SalePrice.Error = SalePrice.Predict- SalePrice,
         SalePrice.AbsError = abs(SalePrice.Predict- SalePrice),
         SalePrice.APE = (abs(SalePrice.Predict- SalePrice)) / SalePrice)%>%
  filter(SalePrice < 5000000)

#creating df just with essential comparison data
bothRegressions <- 
  rbind(
    dplyr::select(boston.test, starts_with("SalePrice"), Regression, Name) %>%
      mutate(lagPriceError = lag.listw(spatialWeights.test, SalePrice.Error)),
    dplyr::select(boston.test.nhood, starts_with("SalePrice"), Regression, Name) %>%
      mutate(lagPriceError = lag.listw(spatialWeights.test, SalePrice.Error)))  

#spatial join of the essential comparison data with our tract data
st_join(bothRegressions, tracts20) %>% 
  group_by(Regression, above75) %>%
  summarize(mean.MAPE = scales::percent(mean(SalePrice.APE, na.rm = T))) %>%
  st_drop_geometry() %>%
  spread(raceContext, mean.MAPE) %>%
  kable(caption = "Test set MAPE by neighborhood racial context")

#calculating lag price error for each model
bothRegressions <- 
  rbind(
    dplyr::select(Charlotte.test, starts_with("price"), Regression, npa) %>%
      mutate(lagPriceError = lag.listw(spatialWeights.test, Price.Error)),
    dplyr::select(Charlotte.test.nhoods, starts_with("price"), Regression, npa) %>%
      mutate(lagPriceError = lag.listw(spatialWeights.test, Price.Error)))   

#table of the abs and ape errors
st_drop_geometry(bothRegressions) %>%
  gather(Variable, Value, -Regression, -npa) %>%
  filter(Variable == "Price.AbsError" | Variable == "Price.APE") %>%
  group_by(Regression, Variable) %>%
    summarize(meanValue = mean(Value, na.rm = T)) %>%
    spread(Variable, meanValue) %>%
    kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

```{r Census Group Analysis Map}

#summarize by MAPE
st_drop_geometry(bothRegressions) %>%
  group_by(Regression, Name) %>%
  summarize(mean.MAPE = mean(SalePrice.APE, na.rm = T)) %>%
  ungroup() %>% 
  left_join(nhoods) %>%
    st_sf() %>%
  #plot the graph
    ggplot() + 
      geom_sf(aes(fill = mean.MAPE)) +
      geom_sf(data = bothRegressions, colour = "black", size = .5) +
      facet_wrap(~Regression) +
      scale_fill_gradient(low = palette5[1], high = palette5[5],
                          name = "MAPE") +
      labs(title = "Mean test set MAPE by neighborhood") +
      mapTheme()

```

## Evaluating our Model

What were some of the more interesting variables? How much of the
variation in prices could you predict? Describe the more important
features? \
Where did the model predict particularly well? Poorly? Why do you think
this might be?

Overall, this is a relatively effective model as it predicts 75% -- 76%
of the home price variation (accounting the spatial variation), with a
mean absolute percentage error (MAPE) at around 22%, both falls in an
acceptable range. We accounted the spatial variation after examining
maps.

### Accuracy

### Generalizability

<Paragraph>

## Conclusion & Recommendation

We would recommend our model to Zillow after further modifications to
the current model. Although the model is already quite effective for
prediction, we see several potential improvements to be made to make it
even more stable and powerful. As we mentioned in the introduction, we
have consulted a local family who are currently into buying house for
some insights, and here is their answer:

"We are actually going to start looking into buying a house soon, so
we\'ve talked recently about these factors. Some of our thoughts have
been:

-   Location seems to be top priority, being in a neighborhood close to
    uptown/SouthEnd (a popular area) is prime real estate

-   Schools would be another consideration (Myers Park and Dilworth are
    sought after)

-   I think more about safety (whether I can go out for a walk/run and
    feel safe)

-   Age of house is not much of a determinant for us, but we have a lot
    of friends that have bought/are buying and want something new, and
    in some of the sought after areas, people have been buying based on
    the land, then putting up new houses

-   We also have a dog, as many people do in Charlotte (very
    dog-friendly city), so finding a place with a yard (preferred
    fenced)

-   Being walking distance to things is also an important factor (we
    loved our apartment we use to live in because we could walk to
    Target and Trader Joe\'s, but needed more space)

    We live in an area called South Park, next to the mall here in
    Charlotte, and love it because one road can take us directly into
    the city, so we don\'t normally take the highway to get anywhere,
    which is really nice."

Although we did take location (neighborhood), schools, safety, year
built, parcel area, and convenient stores into consideration, due to
limited access to free open data, we were not able to engineer factors
that better encapsulate those needs. Hence, some improvements may be:

-   neighborhood as conventionally known instead of administrative
    division of NPA (neighborhood profile area) would be helpful --
    since buyers would look for homes in "South Park" instead of homes
    in "NPA 44"

-   data on schools can be more refined by categorizing primary, middle,
    and high schools, since people don't need 5 primary schools near
    their homes

-   walkability - it would be helpful to have an indicator of
    walkability, not only in terms of distance but also infrastructure
    like safe pedestrian walkways and crossings with traffic lights

-   more indicators related to pet-friendly infrastructure may be
    helpful, too.
