---
title: 'Regrestimate: Predicting Charlotte, NC Home Sale Prices'
author: "Ann (Zi'an) Zhang, Ben Keel"
date: "2022-10-14"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    css:
editor_options: 
  markdown: 
    wrap: 72
---
# Introduction

This project uses the Hedonic Model and OLS regression for predicting home price in Mecklenberg County, NC. We use existing 46,183 cases to build the model -- 60% for training the model and 40% for testing the efficacy of the model. Once we produce a model with satisfactory efficacy, we then use this model to make prediction for 100 home units. 

To generate a model with stronger predictive power, we incorporated more valid internal (structural) and external (neighborhoods, amenities, etc.) factors that may affect home price. The selection and engineering of those factors were executed with careful consideration of Mecklenberg County (including the City of Charlotte)'s local circumstances. After consulting local residents of Charlotte, we included factors like schools, crime, and area of parcel since schooling, safety, and having a yard are among their primary concerns when choosing houses. 

(Brief Results)

## Starting Point
<Describing the set up>


```{r setup, include=FALSE}

# You can set some global options for knitting chunks

knitr::opts_chunk$set(echo = TRUE)

# Load some libraries

library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot) # plot correlation plot
library(corrr)      # another way to plot correlation plot
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance) # to support jtools plots
library(ggpubr)    # plotting R^2 value on ggplot point scatter
library(broom.mixed) # needed for effects plots

# functions and data directory
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#25CB10", "#5AB60C", "#8FA108",   "#C48C04", "#FA7800")
Pink <- c("#ffffff","#f9f4f4","#f0e4e4", "#e7d3d3", "#dec3c3")
Blue <- c("#f8fbff", "#eaf4ff", "#d6eaff", "#add6ff", "#84c1ff")
Violet <- c("#ffffff", "#f7f7f7", "#dfe3ee", "#8b9dc3",  "#58668b")
Green <- c("#e8f4ea", "#e0f0e3", "#d2e7d6","#c8e1cc", "#b8d8be")
Purple <- c("#f3e0f7", "#e4c7f1", "#d1afe8", "#b998dd", "#9f82ce")
Teal <- c("#d1eeea", "#a8dbd9", "#85c4c9", "#68abb8", "#4f90a6")
```

<Describing the original data set>

```{r}

#original dataset
Charlotte.nhoods <- st_read(file.path("D:\\MUSA_508\\Zestimate\\data\\npa"))  %>%
  st_transform('ESRI:103501')

#
Charlotte <-
  st_read("https://raw.githubusercontent.com/mafichman/MUSA_508_Lab/main/Midterm/data/2022/studentData.geojson") %>%
  st_transform('ESRI:103501')


```

## Variables we Liked
<Gathering Data Paragraph>

```{r Variables We Liked}

Charlotte.Clean <-
  Charlotte %>%
  dplyr::select("pid", "nc_pin", "municipali", "yearbuilt", "heatedarea", "price", "storyheigh", "heatedfuel", "actype", "extwall", "foundation", "numfirepla", "fireplaces", "bldggrade", "fullbaths", "halfbaths", "bedrooms", "toPredict", "landusecod", "musaID") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501') %>%
  filter(price <= 10000000)

Charlotte.Clean$fireplaces[is.na(Charlotte.Clean$fireplaces)] <- "FP0"

Charlotte.Clean <-
  st_join(Charlotte.Clean, Charlotte.nhoods, join=st_intersects)

```

Additonal Variables

```{r}

#Park location information
park <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/735a6bce6306442face38657b50fc7b7_10/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_transform('ESRI:103501') 
  
park.buffer <-
  park %>%
  dplyr::select(geometry) %>%
  st_transform('ESRI:103501') %>%
    na.omit() 

Charlotte.Clean$park.Buffer <- 
    Charlotte.Clean$geometry %>% 
    st_buffer(5280) %>% 
    aggregate(mutate(park.buffer, counter = 1),., sum) %>%
    pull(counter)

Charlotte.Clean$park.Buffer[is.na(Charlotte.Clean$park.Buffer)] <- 0


#School location information
school <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/04f2ea0b58774ee7b2e525816cbbc0bb_1/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501') %>%
  filter(!st_is_empty(.))

school.buffer <-
  school %>%
  dplyr::select(geometry) %>%
  st_transform('ESRI:103501') %>%
    na.omit()

Charlotte.Clean$school.Buffer <- 
    Charlotte.Clean$geometry %>% 
    st_buffer(5280) %>% 
    aggregate(mutate(school.buffer, counter = 1),., sum) %>%
    pull(counter)

Charlotte.Clean$school.Buffer[is.na(Charlotte.Clean$school.Buffer)] <- 0


#Hospital Proximity
medical <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/320dbc7d1ef944f5bf7c5e21b018b678_4/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501')

med.buffer <-
  medical %>%
  dplyr::select(geometry) %>%
  st_transform('ESRI:103501') %>%
    na.omit() 

Charlotte.Clean$med.Buffer <- 
    Charlotte.Clean$geometry %>% 
    st_buffer(5280) %>% 
    aggregate(mutate(med.buffer, counter = 1),., sum) %>%
    pull(counter)

Charlotte.Clean$med.Buffer[is.na(Charlotte.Clean$med.Buffer)] <- 0
  

#Food Access
grocery <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/93e54082dde0418d836a57f2fc12879f_7/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_transform('ESRI:103501') %>%
  dplyr::select("OBJECTID", "geometry")

shoppingmall <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/58487298f4ee455e84e236b5db43195d_11/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501') %>%
  dplyr::select("OBJECTID", "geometry")

shopping <-
  rbind(grocery, shoppingmall)

Charlotte.Clean <-
  Charlotte.Clean %>% 
    mutate(
      shop_nn1 = nn_function(st_coordinates(Charlotte.Clean), 
                              st_coordinates(shopping), k = 1))

#Health Access
pharmacies <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/dcfdb72bc9c045a0b0945da79a966841_3/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501') %>%
  filter(!st_is_empty(.))

Charlotte.Clean <-
  Charlotte.Clean %>% 
    mutate(pharm_nn1 = nn_function(st_coordinates(Charlotte.Clean), 
                              st_coordinates(pharmacies), k = 1))

#High-profile crime
homocide <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/b5b21fcd2ad24de9ba7a13093648f5e9_0/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_transform('ESRI:103501')

homicide.buffer <-
  homicide %>%
  dplyr::select(LATITUDE_PUBLIC, LONGITUDE_PUBLIC) %>%
  st_as_sf(coords = c("LONGITUDE_PUBLIC", "LATITUDE_PUBLIC"), crs = 4326) %>%
  st_transform('ESRI:103501') %>%
    na.omit() 

Charlotte.Clean$homi.Buffer <- 
    Charlotte.Clean$geometry %>% 
    st_buffer(5280) %>% 
    aggregate(mutate(homicide.buffer, counter = 1),., sum) %>%
    pull(counter)

Charlotte.Clean$homi.Buffer[is.na(Charlotte.Clean$homi.Buffer)] <- 0


#Low-profile Crime
incidents <- 
  st_read("https://opendata.arcgis.com/api/v3/datasets/d22200cd879248fcb2258e6840bd6726_0/downloads/data?format=geojson&spatialRefId=4326&where=1%3D1") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:103501') %>%
  filter(!st_is_empty(.))

Charlotte.Clean <-
  Charlotte.Clean %>% 
    mutate(inci_nn1 = nn_function(st_coordinates(Charlotte.Clean), 
                              st_coordinates(incidents), k = 1))


```

### Summary Statistics

<Paragraph>
<Data sorted by internal characteristics, local surroundings, spatial structure>

```{r Summary Statistics, warning = FALSE, message = FALSE}



```


### Correlation Matrix

<Paragraph>

```{r Correlation Matrix, warning = FALSE, message = FALSE}

Charlotte.Modelling <-
  Charlotte.Clean %>%
  filter(toPredict=="MODELLING") 

Charlotte.Challenge <-
  Charlotte.Clean %>%
  filter(toPredict=="CHALLENGE")


```

```{r Correlation Matrix, warning = FALSE, message = FALSE, fig.width=10}
numericVars <- 
  select_if(st_drop_geometry(Charlotte.Modelling), is.numeric) %>% na.omit()

ggcorrplot(
  round(cor(numericVars), 1), 
  p.mat = cor_pmat(numericVars),
  colors = c("#B0E0E6", "white", "#B272A6"),
  type="lower",
  insig = "blank") +  
    labs(title = "Correlation across numeric variables") 

```

### Home Price Scatterplots

Present 4 home price correlation scatter plots that you think are of interest.

```{r Home Price Scatterplots, warning = FALSE, message = FALSE,fig.width=6}
#homicide
Charlotte.Modelling %>%  
  st_drop_geometry() %>%
  filter(price <= 1000000) %>%
  dplyr::select(price, starts_with("homo_")) %>%
  gather(Variable, Value, -price) %>%
  ggplot(aes(Value, price)) +
      geom_point(size = .1, color = "light grey") +
      geom_smooth(method = "lm", se=F, color = "black") +
      facet_wrap(~Variable, nrow = 1, scales = "free") +
      labs(title = "Home Price & Distance to the (1st/2nd/3rd) Nearest Homicide", subtitle = "Mecklenberg County, NC") +
      xlab("Distance (in ft.)") +
      ylab("Home Price") +
  theme(strip.text.x = element_text(size = 3)) +
  plotTheme()


```

```{r Home Price Scatterplots, warning = FALSE, message = FALSE, fig.width=6}
#inci_nn1


Charlotte.Modelling %>%  
  st_drop_geometry() %>%
  filter(price <= 1000000) %>%
  dplyr::select(price, starts_with("inci_")) %>%
  gather(Variable, Value, -price) %>%
  ggplot(aes(Value, price)) +
      geom_point(size = .1, color = "#B0C4DE") +
      geom_smooth(method = "lm", se=F, color = "black") +
      facet_wrap(~Variable, nrow = 1, scales = "free") +
      labs(title = "Home Price & Distance to the (1st/2nd/3rd) Nearest Incident", subtitle = "Mecklenberg County, NC") +
      xlab("Distance (in ft.)") +
      ylab("Home Price") +
  theme(strip.text.x = element_text(size = 3)) +
  plotTheme()

```

```{r Home Price Scatterplots, warning = FALSE, message = FALSE, fig.width=6}

#shop_nn

Charlotte.Modelling %>%  
  st_drop_geometry() %>%
  filter(price <= 1000000) %>%
  dplyr::select(price, starts_with("shop_")) %>%
  gather(Variable, Value, -price) %>%
  ggplot(aes(Value, price)) +
      geom_point(size = .1, color = "#D8BFD8") +
      geom_smooth(method = "lm", se=F, color = "dark grey") +
      facet_wrap(~Variable, nrow = 1, scales = "free") +
      labs(title = "Home Price & Distance to the (1st/2nd/3rd) Nearest Shop", subtitle = "Mecklenberg County, NC") +
      xlab("Distance (in ft.)") +
      ylab("Home Price") +
  theme(strip.text.x = element_text(size = 3)) +
  plotTheme()
 
```


```{r Home Price Scatterplots, warning = FALSE, message = FALSE, fig.width=6}

#school_nn
Charlotte.Modelling %>%  
  st_drop_geometry() %>%
  filter(price <= 1000000) %>%
  dplyr::select(price, starts_with("school_")) %>%
  gather(Variable, Value, -price) %>%
  ggplot(aes(Value, price)) +
      geom_point(size = .1, color = "#B4D7BF") +
      geom_smooth(method = "lm", se=F, color = "dark grey") +
      facet_wrap(~Variable, nrow = 1, scales = "free") +
      labs(title = "Home Price & Distance to the (1st/2nd/3rd) School", subtitle = "Mecklenberg County, NC") +
      xlab("Distance (in ft.)") +
      ylab("Home Price") +
  theme(strip.text.x = element_text(size = 3)) +
  plotTheme()


```
### Map of Home Prices and Indicators

<Paragraph>

```{r Map of Home Prices, warning = FALSE, message = FALSE}

ggplot() + 
  geom_sf(data = st_union(Charlotte.nhoods), fill = "#000000") +
  geom_sf(data = Charlotte.Modelling, aes(color = q5(price)), show.legend = "point", size = .1) +
    scale_colour_manual(values = Blue, labels=qBr(Charlotte.Modelling,"price"), name="Home Price") +
    labs(title = "Home Price (Dependent Variable)", subtitle = "Mecklenberg County, NC") +
                          mapTheme()

```

### Map 1 School Buffer


```{r Indicator Map 1, warning = FALSE, message = FALSE}

ggplot() + 
  geom_sf(data = st_union(Charlotte.nhoods), fill = "#000000") +
  geom_sf(data = Charlotte.Modelling, aes(color = q5(school.Buffer)), show.legend = "point", size = .1) +
    scale_colour_manual(values = Pink, labels=qBr(Charlotte.Modelling,"school.Buffer"), name="School Buffer") +
    labs(title = "Number of Schools Within 1 Mile Buffer", subtitle = "Mecklenberg County, NC", aes(color= "pink")) +
                          mapTheme()

```

### Map 2 Shop nn1 

Map 2 of our second most interesting predictor.

```{r Indicator Map 2, warning = FALSE, message = FALSE}

ggplot() + 
  geom_sf(data = st_union(Charlotte.nhoods), fill = "#000000") +
  geom_sf(data = Charlotte.Modelling, aes(color = q5(shop_nn1)), show.legend = "point", size = .1) +
    scale_colour_manual(values = Green, labels=qBr(Charlotte.Modelling,"shop_nn1"), name="Home Price") +
    labs(title = "Distance to the Nearest Shop (Independent Variable 2)", subtitle = "Mecklenberg County, NC") +
                          mapTheme()

```

### Map 3 land use code

Map 3 of our third most interesting predictor.

```{r Indicator Map 3, warning = FALSE, message = FALSE}

ggplot() + 
  geom_sf(data = st_union(Charlotte.nhoods), fill = "#203354") +
  geom_sf(data = Charlotte.Modelling, aes(color=landusecod), show.legend = "point", size = .1) +
    labs(title = "Land Use Code (Inependent Variable)", subtitle = "Mecklenberg County, NC") +
                          mapTheme()

```


# extra engaging stuff

## Prediction Methods

<Paragraph>
OLS Regression

```{r Prediction Methods, warning = FALSE, message = FALSE}

inTrain <- createDataPartition(
              y = paste(Charlotte.Modelling$storyheigh, Charlotte.Modelling$actype, Charlotte.Modelling$aheatingty, Charlotte.Modelling$heatedfuel, Charlotte.Modelling$extwall, Charlotte.Modelling$foundation, Charlotte.Modelling$numfirepla, Charlotte.Modelling$fireplaces, Charlotte.Modelling$bldggrade, Charlotte.Modelling$fullbaths, Charlotte.Modelling$halfbaths, Charlotte.Modelling$bedrooms, Charlotte.Modelling$municipali, Charlotte.Modelling$landusecod), 
              p = .60, list = FALSE)
Charlotte.training <- Charlotte.Modelling[inTrain,] 
Charlotte.test <- Charlotte.Modelling[-inTrain,]  


#Cleaner one
reg.training <- 
  lm(price ~ ., data = as.data.frame(Charlotte.training) %>% 
  dplyr::select(price, yearbuilt, heatedarea, storyheigh, heatedfuel, actype, extwall, foundation, numfirepla, fireplaces, bldggrade, fullbaths, halfbaths, bedrooms, pharm_nn1, inci_nn1, homo.Buffer, school.Buffer, med.Buffer, municipali, landusecod, park.Buffer, shop_nn1, shape_Area.x))
  
```

## Results

### Training Set Results

<Paragraph>

```{r Training Set, warning = FALSE, message = FALSE}

summary(reg.training)

Charlotte.test <-
  Charlotte.test %>%
  mutate(Regression = "Baseline Regression",
         Price.Predict = predict(reg.training, Charlotte.test),
         Price.Error = Price.Predict - price,
         Price.AbsError = abs(Price.Predict - price),
         Price.APE = (abs(Price.Predict - price)) / Price.Predict) %>%
  filter(price < 10000000) %>%
  na.omit(Charlotte.test)


mean(Charlotte.test$Price.APE, na.rm = T)

```

### Test Set Results

<Paragraph>

```{r Test Set Results, warning = FALSE, message = FALSE}



```

### Cross-Validation Tests & Results

<Paragraph>

```{r CV Results, warning = FALSE, message = FALSE}



```

### Predicted vs Observation Scatterplot

<Paragraph>

```{r Predicted vs Observation Scatterplot, warning = FALSE, message = FALSE}



```

### Residuals Map

<Paragraph>

```{r Residuals Map, warning = FALSE, message = FALSE}



```

### Predicted Map

<Paragraph>

```{r Predicted Map, warning = FALSE, message = FALSE}



```

### MAPE by Neighborhood

<Provide a polished table of mean absolute error and MAPE for a single test set.>

<Paragraph> 

```{r MAPE by Neighborhood, warning = FALSE, message = FALSE}

reg.nhood <- lm(price ~ ., data = as.data.frame(Charlotte.training) %>% 
                                 dplyr::select(npa, price, yearbuilt, heatedarea, 
                                               storyheigh, heatedfuel, actype, extwall, 
                                               foundation, numfirepla, fireplaces, 
                                               bldggrade, fullbaths, halfbaths, bedrooms, 
                                               park_nn3, pharm_nn1, inci_nn1, 
                                               homo.Buffer, school.Buffer, med.Buffer, municipali, landusecod))

summary(reg.nhood)


Charlotte.test.nhoods <-
  Charlotte.test %>%
  mutate(Regression = "Neighborhood Effects",
         Price.Predict = predict(reg.nhood, Charlotte.test),
         Price.Error = Price.Predict - price,
         Price.AbsError = abs(Price.Predict - price),
         Price.APE = (abs(Price.Predict - price)) / Price.Predict) %>%
  filter(price < 10000000) 



bothRegressions <- 
  rbind(
    dplyr::select(Charlotte.test, starts_with("price"), Regression, npa) %>%
      mutate(lagPriceError = lag.listw(spatialWeights.test, Price.Error)),
    dplyr::select(Charlotte.test.nhoods, starts_with("price"), Regression, npa) %>%
      mutate(lagPriceError = lag.listw(spatialWeights.test, Price.Error)))   

st_drop_geometry(bothRegressions) %>%
  gather(Variable, Value, -Regression, -npa) %>%
  filter(Variable == "Price.AbsError" | Variable == "Price.APE") %>%
  group_by(Regression, Variable) %>%
    summarize(meanValue = mean(Value, na.rm = T)) %>%
    spread(Variable, meanValue) %>%
    kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

#### Scatterplot - MAPE by neighborhood mean price

<Provide a scatterplot plot of MAPE by neighborhood as a function of mean price by neighborhood.>

<Paragraph>

```{r MAPE scatterplot, warning = FALSE, message = FALSE}



```

### Split by Census Group

<Using tidycensus, split your study area into two groups (perhaps by race or income) and test your model’s generalizability. Is your model generalizable?>

<Paragraph>

```{r Split by Census Group, warning = FALSE, message = FALSE}



```

## Evaluating our Model

Is this an effective model? 
What were some of the more interesting variables?
How much of the variation in prices could you predict?
Describe the more important features? Describe the error in your predictions? 
According to your maps, could you account the spatial variation in prices?  
Where did the model predict particularly well? Poorly? Why do you think this might be?

### Accuracy


### Generalizability

<Paragraph>

## Conclusion